Democracy, Public Support, and Measurement Uncertainty

### 2020-12-31 paper/dcpo_demsupport.Rmd
rewrite first paragraph for balance; minimalist title per https://github.com/fsolt/dcpo_dem_mood/issues/11#issuecomment-752794638

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Public Support for Democracy, Regime Change, and the Importance of Incorporating Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehu@mail.tsinghua.edu.cn](mailto:yuehu@mail.tsinghua.edu.cn).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in regime? Two prominent recent studies (Claassen 2020_a_, Claassen 2020_b_) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this uncertainty into its analyses. Here we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis. [142/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---
# Introduction

One classic but disputable question in comparative politics is the relationship between mass democratic support and countries' level of democracy.
A long-standing and widely accepted theory is that the level of mass democratic support determines the stability of a democratic system [e.g., @lipset_social_1959; @easton1965systems; @norris_democratic_2011], but empirical findings remain inconclusive [@inglehart_modernization_2005; @fails_changing_2010; @qi_how_2011; for a review of this literature, see @Claassen2020a].

One reason for the mixed results is the difficulties in measuring comparative public democratic support.
Public support for democracy is a latent variable, and the incorrect measurement of it may limit inferences about relationship between public opinion and democratic institutional development.
However, the sparsity and incomparability of public opinion data across countries over time have always been perplexing comparative studies for a long time [@Solt2020a].
Recently, a few pioneering studies have contributed to comparative public opinion studies using advanced latent variable measurement [@Caughey2019; @Claassen2019; @Solt2020a].

A prominent study by @Claassen2020a applied an advanced Bayesian Item Response Model to measure dynamic public democratic support for 135 countries for up to 29 years.
It found that mass support had a positive impact on democratic change, especially the endurance of democracy.
The analysis models he used do not account, though, for measurement uncertainty.
The latent constructs are inherently accompanied by a quantity  of measurement uncertainty due to the unobservability of variables.
Left unacknowledged, with a strong assumption that latent variables are measured perfectly, measurement error could distort both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

This paper reexamined the classic claim about public democratic attitude and democratic changes thoroughly through both sophisticated methods and bigger data.
In addition to incorporating uncertainty, we employed a more sophisticated measurement, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, to estimate democratic support for a larger dataset with 144 countries for up to 33 years between 1988 and 2020.
Surprisingly, our study revealed that the significant relationship between public support and democratic change disappeared once uncertainty was included in the replication analysis of @Claassen2020a model with his original data and the DCPO analysis with more data.
To be cautious, we then also reexamined @Claassen2020b analyses by following the same operations. 
Again, when uncertainty was incorporated, the significant relationship between democratic changes and public democratic attitude disappeared.

The disappearing relationship might be caused by theoretical, methodological reasons, or both.
Methodologically, the existing survey questions might not be able to capture mass democratic attitude due to a shortage of indicators [Solt, 2021??] and the ineffectiveness of indicators in measuring the multi-faceted democratic support [@wuttke2020have] and behaviorally democratic commitments [@graham_democracy_2020; @Carey2020; @McCoy2020].
Substantive explanations might also account for the insignificant relationship: the elite-driven democratic backsliding [@weingast_political_1997; @levitsky_how_2018], the lack of a consensus on democratic erosion between the public and experts [@carey_searching_2019], and the difference between the public support in abstract and public commitments in context [@bartels_ethnic_2020; @graham_democracy_2020; @McCoy2020;@prothro_fundamental_1960;@van2017myth
].

>>>>

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Democracy, Public Support, and Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehu@mail.tsinghua.edu.cn](mailto:yuehu@mail.tsinghua.edu.cn).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in democracy? Two prominent recent studies (Claassen 2020a, Claassen 2020b) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this uncertainty into its analyses. Here we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis. [142/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---
# Introduction

It has long been argued that democratic regimes and public support for them are mutually reinforcing: that high levels of public support ensure democracies remain strong, and that experience with democratic governance generates robust public support [see, e.g., @lipset_social_1959; @Easton1965].
But the evidence for either side of this claim has been decidedly mixed.
Countries with greater democratic support have been found to become stronger and more stable democracies [e.g., @Inglehart2005, 251-254] and just the opposite [@Fails2010, 182-183].
Similarly, studies have alternately found that more experience with democracy yields more democratic support [e.g., @Fails2010, 183; @Wuttke2020a, 5-6] or instead that long-established democracies are suffering from democratic fatigue [e.g., @Denemark2016; @Foa2017].

One important reason for these mixed results is the difficulty in measuring democratic support over time and across many countries.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationships between public opinion and institutional development.
Further, the survey data available across countries and over time on support for democracy--or indeed most topics in public opinion---are sparse and incomparable, greatly hindering the broadly comparative research needed to reach sound conclusions.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data and advance broadly comparative research by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
A pair of prominent recent works took advantage of this latent variable approach to measure dynamic public democratic support for 135 countries for up to 29 years and to then assess its consequences for democratic change.
It concluded that mass support had a positive impact on democratic change, especially the endurance of democracy.

The models employed in this study's analyses, though, do not account for measurement uncertainty.
Because they are unobserved, latent variables are inherently accompanied by a quantity of measurement uncertainty.
To leave this uncertainty unacknowledged is to make the implausible assumption that the latent variables are measured perfectly, an assumption which distorts both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

This paper reexamines the classic claim about public democratic attitude and democratic changes thoroughly through both sophisticated methods and bigger data.
In addition to incorporating uncertainty, we employed a more sophisticated measurement, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, to estimate democratic support for a larger dataset with 144 countries for up to 33 years between 1988 and 2020.
Surprisingly, our study revealed that the significant relationship between public support and democratic change disappeared once uncertainty was included in the replication analysis of @Claassen2020a model with his original data and the DCPO analysis with more data.
To be cautious, we then also reexamined @Claassen2020b analyses by following the same operations. 
Again, when uncertainty was incorporated, the significant relationship between democratic changes and public democratic attitude disappeared.

The disappearing relationship might be caused by theoretical, methodological reasons, or both.
Methodologically, the existing survey questions might not be able to capture mass democratic attitude due to a shortage of indicators [Solt, 2021??] and the ineffectiveness of indicators in measuring the multi-faceted democratic support [@wuttke2020have] and behaviorally democratic commitments [@graham_democracy_2020; @Carey2020; @McCoy2020].
Substantive explanations might also account for the insignificant relationship: the elite-driven democratic backsliding [@weingast_political_1997; @levitsky_how_2018], the lack of a consensus on democratic erosion between the public and experts [@carey_searching_2019], and the difference between the public support in abstract and public commitments in context [@bartels_ethnic_2020; @graham_democracy_2020; @McCoy2020; @prothro_fundamental_1960; @van2017myth
].
